Variational Monte Carlo on bosonic systems

### 2018-03-22 project_1/article/article.tex
Wrote about blocking.

@@

        		% Defining blocks of flowchart
        		\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
   			 	text width=5em, text badly centered, node distance=3cm, inner sep=0pt]
			\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    				text width=10em, text centered, rounded corners, minimum height=4em]
			\tikzstyle{line} = [draw, -latex']

@@

    				% Nodes
    				\node [block] (init) {Start with $\mathbf{X}_i$, $i=0$};
    				\node [block, below of=init] (variance) {Compute variance and autocovariance of $\mathbf{X}_i$, $s_i$ and $\gamma_i$};
    				\node [block, below of=variance] (transform) 
    					{Transform data $(\mathbf{X}_{i+1})_k = \frac{1}{2}\left((\mathbf{X})_i)_{2k-1} + (\mathbf{X}_i)_{2k} \right)$};
    				\node [block, right=1.5cm of transform] (i+1) {Next iteration: $i = i + 1$.} ;
    				\node [decision, below of = transform] (x_size) {Length of $\mathbf{X}_{i + 1} \geq 2$ ? }; 
    				\node [block, below of=x_size] (m_test) {Compute test statistic $M_j$};
    				\node [block, below of=m_test] (magic_numbers) {Find first $k$ so that $M_k \leq q_{d-k}(1-\alpha)$};
    				\node [block, below of=magic_numbers] (return) {Return $s_k^2/n_k$};

    				% Edges
    				\path [line] (init) -- (variance);
    				\path [line] (variance) -- (transform);
    				\path [line] (transform) -- (x_size);
    					\path [line] (x_size) -| node [near start] {yes} (i+1);
    					\path [line] (i+1) |- (variance);
    				\path [line] (x_size) -- node [near start] {no} (m_test);
    				\path [line] (m_test) -- (magic_numbers);
    				\path [line] (magic_numbers) -- (return);
    			\end{tikzpicture}
    			\caption{Flow chart, blocking algorithm.}
		\end{figure}

>>>>

     	Instead of going by this "manual" method of looking at charts, we will construct a test statistic and
     	let the computer do the necessary considerations in a more "automatic way.

        The easiest way to do this is by use of the \emph{automatic blocking scheme}, nicely illustrated 
        with a flow chart in \autoref{fig:blocking}. The input parameter for the method is an array of the 
        local energies, but it can be any array of time series data, which is why it is herein referred to as
        $\mathbf{X}$. Any array of sequential data, where any particular data point depends on the previous
        data point inhibits all attributes of a time series. Our array of local energies is therefor most definitely 
        a time series.

		From $\mathbf{X}$ we compute the variance and sample covariance between the series itself and a 
		one-point lag of the series. This covariance is usually referred to as the first order autocovariance.
		\begin{equation}
			\gamma_i = \frac{1}{n}
				\left(\sum_{k=1}^n\left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)\right)
				\left(\sum_{k=0}^{n-1}\left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)\right).
		\end{equation}

		Now we "block"! The next step is to perform a transformation of the our time series $\mathbf{x}_i$, where
		we end up with an array that is half the length of the original time series,
		\begin{equation}
			(\mathbf{X}_i)_k = \frac{1}{2}\left((\mathbf{X}_i)_{2k-1} + (\mathbf{X}_i)_{2k} \right).
		\end{equation}
		This process is were the term "blocking" stems from. There are a number of ways to perform the blocking -
		we could have picked the data points used in the new ($i+1$) array randomly, or more orderly. Here we we have
		picked them sequentially. We have obtained $\frac{n}{2}$ new "primed" stochastic variables. Assuming that the
		reader is playing close attention, it is now needless to say that it is much easier to perform these sequential
		blocking steps if we have an array of length $n= 2^d$, where $d$ is integer. Otherwise, problems would arise
		because one would have arrays of different lengths when the original array is "split in two". The ailment is
		easy to remedy by excluding an observation from the longest resulting array after splitting.

		For each new time series $\mathbf{X}_{i+1}$ a new sample variance and sample covariance is computed.
		The blocking transformation is allowed to continue until we are we run out of data, i.e. the length
		of $\mathbf{X}_i < 2$. In other words, there is only a single number left. We now compute a test statistic,
		\begin{equation}
			M_j = \sum_{k=j}^{d-1} n_k \frac{\left[(n_k - 1)\frac{s^2_k}{n_k^2} y \gamma_k(1)\right] ^2}{s_k^4}.
		\end{equation}
		As luck would have it, this number $M$ is   $\chi^2$-distributed. The following step would be to look up a 
		statistical table and find the first $k$, such that $M_k \leq q_{d-k}(1-\alpha)$, where $q$ is the statistical
		limit and $\alpha$ is  the significance level that your hear desires. Finally, the correct sample standard deviation 
		is $\frac{s_k^2}{n_k}$. Again, see \autoref{fig:blocking} for a summary of the algorithm.

@@

        		% Defining blocks of flowchart
        		\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
   			 	text width=5em, text badly centered, node distance=3cm, inner sep=0pt]
			\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    				text width=11em, text centered, rounded corners, minimum height=4em]
			\tikzstyle{line} = [draw, -latex']

@@

    				% Nodes
    				\node [block] (init) {Start with $\mathbf{X}_i$, $i=0$};
    				\node [block, below of=init] (variance) {Compute variance and autocovariance of $\mathbf{X}_i$, $s_i$ and $\gamma_i$};
    				\node [block, below of=variance] (transform) 
    					{Transform data $(\mathbf{X}_{i+1})_k = \frac{1}{2}\left((\mathbf{X})_i)_{2k-1} + (\mathbf{X}_i)_{2k} \right)$};
    				\node [block, right=1.5cm of transform] (i+1) {Next iteration: $i = i + 1$.} ;
    				\node [decision, below of = transform] (x_size) {Length of $\mathbf{X}_{i + 1} \geq 2$ ? }; 
    				\node [block, below of=x_size] (m_test) {Compute test statistic $M_j$ from $s_j$ and $\gamma_j$ for all $j$.};
    				\node [block, below of=m_test] (magic_numbers) {Find first $k$ so that $M_k \leq q_{d-k}(1-\alpha)$};
    				\node [block, below of=magic_numbers] (return) {Return $s_k^2/n_k$};

    				% Edges
    				\path [line] (init) -- (variance);
    				\path [line] (variance) -- (transform);
    				\path [line] (transform) -- (x_size);
    					\path [line] (x_size) -| node [near start] {yes} (i+1);
    					\path [line] (i+1) |- (variance);
    				\path [line] (x_size) -- node [near start] {no} (m_test);
    				\path [line] (m_test) -- (magic_numbers);
    				\path [line] (magic_numbers) -- (return);
    			\end{tikzpicture}
    			\caption{Flow chart depicting the automatic blocking method for finding a better estimate for sample variance.}
    			\label{fig:blocking}
		\end{figure}
